#This is the part of the project that I specifically worked on. I analyzed which starters were good in the first 15 pitches of an outing compared to the rest of their outing, looking for pitchers who would start well and fall off as time passed because those players would likely fare better in the bullpen.

---
title: "Reds"
author: "Colin Hofmeister"
date: "2024-01-22"
output: html_document
---
```{r}

library(tidyverse)
library(ggplot2)
library(ggrepel)
setwd("C:/Users/Colin/OneDrive/Pictures/2024 Hackathon")
savant <- read_csv("savant_pitch_level.csv")
fangraphs <- read_csv("fangraphs_season_level.csv")
```



```{r}
starters <- savant %>% filter(sp_indicator==1) %>% group_by(player_name, pitcher) %>% summarise(num_pitches = n())

starters <- starters[starters$num_pitches > 1000,]

qualified <- savant  %>% filter(pitcher %in% starters$pitcher)
```

```{r}
first_15 <- qualified %>% filter(pitch_number_appearance < 15)
rest <- qualified %>% filter(pitch_number_appearance > 15)

first_15 <- first_15 %>% group_by(pitcher,player_name) %>% 
  summarise(first15_woba = sum(woba_value,na.rm = TRUE)/ sum(woba_denom,na.rm = TRUE))

rest <- rest %>% group_by(pitcher,player_name) %>% 
  summarise(rest_woba = sum(woba_value,na.rm = TRUE)/ sum(woba_denom,na.rm = TRUE))

total <- merge(first_15,rest, by=c("pitcher","player_name"))

total <- total %>% 
  mutate(rest_sub_first15_woba = rest_woba-first15_woba) %>% 
  arrange(-rest_sub_first15_woba)
top <- head(total, 100)

top$player_name <- reorder(top$player_name, -top$rest_sub_first15_woba)

write.csv(top, "top_first_15.csv")

avg <- total %>% 
  filter(!is.na(first15_woba)) %>% 
  mutate(first15_woba = as.numeric(first15_woba))

  F15mean_woba = mean(total$first15_woba)
  F15mean_woba

avgr <- total %>% 
  filter(!is.na(rest_woba)) %>% 
  mutate(rest_woba = as.numeric(rest_woba))

  Rmean_woba = mean(total$rest_woba)
  Rmean_woba
```

average first15_woba = 0.338629879
average rest_woba = 0.330754093

```{r}
ggplot() +
  geom_point(data = total, aes(x = first15_woba, y = rest_woba)) +
  geom_text_repel(data = subset(total, 
                               abs(scale(first15_woba)) > 2 | abs(scale(rest_woba)) > 2.5),
                  aes(x = first15_woba, y = rest_woba, label = player_name),
                  vjust = -0.3, hjust = 0.3) + 
  labs(x = "wOBA First 15 Pitches", 
       y = "wOBA After 15 Pitches", 
       title = "Scatter Plot of wOBA First 15 Pitches of an Outing vs Rest of the Outing") +
  theme_minimal() +
  geom_hline(yintercept = .320, linetype = "dashed", color = "black") +
  geom_vline(xintercept = .320, linetype = "dashed", color = "black") 

```



